---
title: "Principal components analysis of 1000 Genomes data"
author: "Peter Carbonetto"
output:
  html_document:
    theme: readable
    include:
      before_body: header.html
      after_body: footer.html
---

This is a small examlpe to illustrate how principal components analysis
(PCA) can be applied to genotype data---specifically, genotype data
from the [1000 Genomes Project][1kg]---to gain insight into population
structure. Here the data preparation steps are skipped.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment   = "#",
                      collapse  = TRUE,
                      fig.align = "center")
```

## Set up environment

Load the [ggplot2][ggplot2], [cowplot][cowplot] and [rsvd][rsvd]
packages. The rsvd package is used to compute the principal components
from the genotype data. For repeatable results, we initialize the
sequence of pseudorandom numbers beacuse the "rpca" function is based
on a randomized algorithm for the singular value decomposition (SVD).

```{r setup, message=FALSE}
library(rsvd)
library(ggplot2)
library(cowplot)
set.seed(1)
```

# Session info

[1kg]: https://doi.org/10.1038/nature15393
[ggplot2]: https://cran.r-project.org/package=ggplot2
[cowplot]: https://cran.r-project.org/package=cowplot
[rsvd]: https://cran.r-project.org/package=rsvd
